Preparación de datos

### 2022-06-13 production/prod_prep.Rmd
analysis

@@

@@

elsoc_long_2016_2021$quintil1 <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$quintil1,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(elsoc_long_2016_2021$quintil1)
#Age Ranges
elsoc_long_2016_2021$edad <- 
  factor(car::recode(elsoc_long_2016_2021$m0_edad, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
elsoc_long_2016_2021$edad <-
  sjlabelled::set_label(elsoc_long_2016_2021$edad, 
                        label = c("Age groups")) 
# Other controls______________________________________________________________
# Subjective social status: individual ____________________________

elsoc_long_2016_2021$ess <- 
as.numeric(elsoc_long_2016_2021$d01_01)

sjmisc::frq(elsoc_long_2016_2021$ess)
elsoc_long_2016_2021$ess <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$ess,
                      label = "Subjective Social Status: individual")

@@

elsoc$nation_t1 <- set_label(elsoc$nation_t1, label = "Nationality")
elsoc$nation_t1 <- set_labels(elsoc$nation_t1,
                              labels =c("Chilean","Peruvian",
                                        "Argentina","Colombian",
                                        "Bolivian", "Haitian","Spanish","Other"))
elsoc$nation_t1 <- factor(elsoc$nation_t1,
                          levels = 1:8,
                          labels = c("Chilean","Peruvian",
                                     "Argentina","Colombian",
                                     "Bolivian", "Haitian","Spanish","Other"))
sjmisc::frq(elsoc$nation_t1)
# KNOW inmigrants _____________________________
sjPlot::tab_xtab(elsoc$know_inm,
                 elsoc$ola,show.cell.prc = T,show.na = T)
sjmisc::frq(elsoc$know_inm)
elsoc$know_inm_bi <- car::recode(elsoc$know_inm,
                                                "1=0;2:5=1")
sjmisc::frq(elsoc$know_inm_bi)

@@

elsoc <- left_join(x = elsoc,y = know_inm_bi, by = "idencuesta")
sjmisc::frq(elsoc$know_inm_t1t4)
elsoc$know_inm_t1t4 <- set_label(elsoc$know_inm_t1t4,
                                  label = "Know immigrants (diff. t4 - t1)")
# sjmisc::frq(elsoc$know_inm_t1t4)
# sjPlot::tab_xtab(elsoc$know_inm_t1t4,elsoc$ola)

@@

elsoc <- left_join(x = elsoc,y = frien_inm_bi, by = "idencuesta")
summary(elsoc$frien_inm_t1t4)
# if NA in difference, impute data from t4
elsoc$frien_inm_t1t4 <- if_else(condition = is.na(elsoc$frien_inm_t1t4),
                                true = elsoc$frien_inm_bi_t4,
                                false = elsoc$frien_inm_t1t4 )
summary(elsoc$frien_inm_t1t4)
# if NA in difference, impute data from t1
elsoc$frien_inm_t1t4 <- if_else(condition = is.na(elsoc$frien_inm_t1t4),
                                true = elsoc$frien_inm_bi_t1,
                                false = elsoc$frien_inm_t1t4)
summary(elsoc$frien_inm_t1t4)

@@

# CASEN nationality ___________________________________________________________
# 2017_________________________________________________________
df_17 <- 
casen2017 %>% 
  select(comuna,starts_with("r1b"), starts_with("exp"))
# sjmisc::frq(df_17$r1b)

#dummy for inmigrant 
df_17$r1b[df_17$r1b==9] <- NA
# sjmisc::frq(df_17$r1b)
df_17$inm <- ifelse(test = (df_17$r1b==3),yes = 1,no = 0)
# sjmisc::frq(df_17$inm)

@@

summary(df_17_comuna$inm_2017_cas)  #check
summary(df_17_comuna$inm_2017_cas_w)#check
# 2020_________________________________________________________
df_20 <- 
casen20 %>% 
  select(comuna,
         starts_with("r1b"),starts_with("exp"))

#dummy for inmigrant 
df_20$r1b[df_20$r1b==9] <- NA
# sjmisc::frq(df_20$r1b)
df_20$inm <- ifelse(test = (df_20$r1b==3),yes = 1,no = 0)
# sjmisc::frq(df_20$inm)

@@

# Label data____________________________________________________________________
sjlabelled::set_label(df_study1_comunas) <- 
c("Comuna",
  "Codigo Comuna",
  "Proporcion de poblacion inmigrante  (ponderado) - CASEN 2017 - Comunal",
  "Número de poblacion inmigrante - CENSO 2017 - Comunal",
  "Proporcion de poblacion inmigrante - CENSO 2017 - Comunal",
  "Número de poblacion inmigrante - CIT 2017 - Comunal (usando zona censal)",
  "Proporcion de poblacion inmigrante (CIT / DEM)- CIT 2017 y INE/DEM 2017 - Comunal",
  "Proporcion de poblacion inmigrante proyectada -  INE/DEM 2018 - Comunal",
  "Proporcion de poblacion inmigrante proyectada -  INE/DEM 2019 - Comunal",
  "Proporcion de poblacion inmigrante proyectada -  INE/DEM 2020 - Comunal",
  "Proporcion de poblacion inmigrante (ponderado) - CASEN 2020 - Comunal",
  "Cambio en la proporción de población inmigrante entre 2017 y 2020 - CASEN")

>>>>

#reverse education, reference level is the highest level
elsoc_long_2016_2021$educ <- forcats::fct_rev(elsoc_long_2016_2021$educ)

@@

#reverse quintile, reference level is the highest quintile
elsoc_long_2016_2021$quintil <- forcats::fct_rev(elsoc_long_2016_2021$quintil)

@@

elsoc_long_2016_2021$quintil1 <- factor(elsoc_long_2016_2021$quintil1, c("Q5","Q4","Q3","Q2","Q1","QNA"))

elsoc_long_2016_2021$quintil1 <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$quintil1,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(elsoc_long_2016_2021$quintil1)

#Age Ranges
elsoc_long_2016_2021$edad <- 
  factor(car::recode(elsoc_long_2016_2021$m0_edad, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
elsoc_long_2016_2021$edad <-
  sjlabelled::set_label(elsoc_long_2016_2021$edad, 
                        label = c("Age groups")) 
# Other controls______________________________________________________________
# Subjective social status: individual ____________________________

elsoc_long_2016_2021$ess <- 
as.numeric(elsoc_long_2016_2021$d01_01)
sjmisc::frq(elsoc_long_2016_2021$ess)
elsoc_long_2016_2021$ess <- sjmisc::rec(elsoc_long_2016_2021$ess,rec = "rev")
sjmisc::frq(elsoc_long_2016_2021$ess)

elsoc_long_2016_2021$ess <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$ess,
                      label = "Subjective Social Status: individual (reverse)")

@@

elsoc$nation_t1 <- set_labels(elsoc$nation_t1,
                              labels =c("Chilean","Peruvian",
                                        "Argentina","Colombian",
                                        "Bolivian", "Haitian","Spanish","Other"))
elsoc$nation_t1 <- factor(elsoc$nation_t1,
                          levels = 1:8,
                          labels = c("Chilean","Peruvian",
                                     "Argentina","Colombian",
                                     "Bolivian", "Haitian","Spanish","Other"))
elsoc$nation_t1 <- set_label(elsoc$nation_t1, label = "Nationality")
sjmisc::frq(elsoc$nation_t1)
# KNOW inmigrants _____________________________
sjPlot::tab_xtab(elsoc$know_inm,
                 elsoc$ola,show.cell.prc = T,show.na = T)
sjmisc::frq(elsoc$know_inm)
elsoc$know_inm_bi <- car::recode(elsoc$know_inm,
                                                "1=0;2:5=1")
sjmisc::frq(elsoc$know_inm_bi)

@@

elsoc <- left_join(x = elsoc,y = know_inm_bi, by = "idencuesta")
sjmisc::frq(elsoc$know_inm_t1t4)

elsoc$know_inm_t1t4 <- factor(elsoc$know_inm_t1t4,levels = c(-1,0,1),
                              labels = c("No longer know","Still know","Now know"))

elsoc$know_inm_t1t4 <- set_label(elsoc$know_inm_t1t4,
                                  label = "Know immigrants (diff. t4 - t1)")
sjmisc::frq(elsoc$know_inm_t1t4)
# sjPlot::tab_xtab(elsoc$know_inm_t1t4,elsoc$ola)

@@

elsoc <- left_join(x = elsoc,y = frien_inm_bi, by = "idencuesta")
summary(elsoc$frien_inm_t1t4)
# if NA in difference, impute data from t4
elsoc$frien_inm_t1t4 <- if_else(condition = is.na(elsoc$frien_inm_t1t4),
                                true = elsoc$frien_inm_bi_t4,
                                false = elsoc$frien_inm_t1t4 )
summary(elsoc$frien_inm_t1t4)
# if NA in difference, impute data from t1
elsoc$frien_inm_t1t4 <- if_else(condition = is.na(elsoc$frien_inm_t1t4),
                                true = elsoc$frien_inm_bi_t1,
                                false = elsoc$frien_inm_t1t4)
sjmisc::frq(elsoc$frien_inm_t1t4)

elsoc$frien_inm_t1t4 <- factor(elsoc$frien_inm_t1t4,
                               levels = c(-1,0,1),
                               labels = c("No longer have","Still have","Now have"))

@@

# CASEN nationality ___________________________________________________________
# 2017_________________________________________________________
df_17 <- 
casen2017 %>% 
  select(comuna,starts_with("r1"), starts_with("exp"))
sjmisc::frq(df_17$r1a)
sjmisc::frq(df_17$r1a_cod)
sjmisc::frq(df_17$r1b)
sjmisc::frq(df_17$r1b_p_cod)

# Latam & caribean 401:588
# recent migrant groups (2020) 
  # Venezuela: 513
  # Peru: 509
  # Haiti: 413
  # Colombia: 505
  # Bolivia: 502

# Al igual que la tendencia reflejada en las estimaciones de 2018 y 2019, la mayoría (poco más de 79% del total) de las personas extranjeras residentes en el país al 31 de diciembre de 2020 proviene de Venezuela (30,7%), Perú (16,3%), Haití (12,5%), Colombia (11,4%) y Bolivia (8,5%).

#dummy for inmigrant 
df_17$pais <- 
car::recode(df_17$r1b_p_cod,
            recodes = "c(513,509,413,505,502)=1;c(888,999)=NA;else=0",
            as.factor = T)
sjmisc::frq(df_17$pais)

# df_17$r1b[df_17$r1b==9] <- NA
# sjmisc::frq(df_17$r1b)

df_17$inm <- ifelse(test = (df_17$pais==1),yes = 1,no = 0)

sjmisc::frq(df_17$inm)

@@

summary(df_17_comuna$inm_2017_cas)  #check
summary(df_17_comuna$inm_2017_cas_w)#check
# 2020_________________________________________________________
df_20 <- 
casen20 %>% 
  select(comuna,
         starts_with("r1"),starts_with("exp"))

sjmisc::frq(df_20$r1b_p_cod)
#dummy for inmigrant 
df_20$pais <- 
car::recode(df_20$r1b_p_cod,
            recodes = "c(513,509,413,505,502)=1;c(888,999)=NA;else=0",
            as.factor = T)
sjmisc::frq(df_20$pais)

# df_20$r1b[df_20$r1b==9] <- NA
# sjmisc::frq(df_20$r1b)
df_20$inm <- ifelse(test = (df_20$pais==1),yes = 1,no = 0)
# sjmisc::frq(df_20$inm)

@@

# Label data____________________________________________________________________
sjlabelled::set_label(df_study1_comunas) <- 
c("Comuna",
  "Codigo Comuna",
  "Proportion of immigrant population (weighted) - CASEN 2017 - Municipality",
  "Number of immigrant population - CENSUS 2017 - Municipality",
  "Proportion of immigrant population - CENSUS 2017 - Municipality",
  "Number of immigrant population - CIT 2017 - Municipality (using census tract)",
  "Proportion of immigrant population (CIT / DEM)- CIT 2017 and INE/DEM 2017 - Communal",
  "Proportion of projected immigrant population - INE/DEM 2018 - Communal",
  "Proportion of projected immigrant population -  INE/DEM 2019 - Comunal",
  "Proportion of projected immigrant population -  INE/DEM 2020 - Comunal",
  "Proportion of immigrant population (weighted) - CASEN 2020 - Municipality",
  "Change in the proportion of the immigrant population between 2017 and 2020 - CASEN")
